---
title: "Model misspecification, a busulfan case study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Busulfan MIPD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, 
  message = FALSE
)
```

In this vignette, we will demonstrate the following tools:

* using {mipdtrial} to simulate a trial, with the trial design specified using
  R code.
* simulating patient PK and estimating patient PK with different models, to
  assess the impact of model misspecification
* using a weight-based initial dose

To account for model misspecification, we will use one model to simulate
(McCune, 2014) and one model to estimate (Shukla, 2020). This is a simplified
version of a [published
study](https://link.springer.com/article/10.1007/s10928-024-09915-w) in which we
compared non-compartmental analysis (NCA) with MAP Bayesian estimation for
AUC-guided busulfan dosing.

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# for CI, most windows users should not need this since the .libPaths should
# include where the model was installed.
if (.Platform$OS.type == "windows") { 
    .libPaths(c("D:/a/mipdtrial/mipdtrial/check/mipdtrial.Rcheck", .libPaths()))
}
if(!require(pkbusulfanmccune, warn.conflicts = FALSE)) {
  capture.output(PKPDsim::install_default_literature_model("pk_busulfan_mccune"))
}
if (!require(pkbusulfanshukla, warn.conflicts = FALSE)) {
  capture.output(capture.output(PKPDsim::install_default_literature_model("pk_busulfan_shukla")))
}
```

```{r setup, message=FALSE, warning=FALSE}
library(mipdtrial)
if(!require(pkbusulfanmccune)) {
  PKPDsim::install_default_literature_model("pk_busulfan_mccune")
  library(pkbusulfanmccune)
}
if (!require(pkbusulfanshukla)) {
  PKPDsim::install_default_literature_model("pk_busulfan_shukla")
  library(pkbusulfanshukla)
}

# For example data
library(NHANES)

# For data handling/plotting
library(dplyr)
library(tidyr)
library(ggplot2)

```

## 1. Define trial design

In this simple trial, patients will be dosed once per day for four days, with 4
levels collected per day, at 3.5, 4, 6 and 8 hours post-dose. We will adjust
doses 2, 3, and 4 to achieve a target cumulative AUC of 90 by t = 192 hours.

```{r}
tdm_design <- create_sampling_design(
  offset = rep(c(3.5, 4, 6, 8), 4),
  at = rep(1:4, each = 4),
  anchor = "dose"
)
target_design <- create_target_design(
  targettype = "cum_auc", 
  targetvalue = 90 * 1000,       # unit conversion
  time = 192
)
update_design <- create_regimen_update_design(
  at = c(2, 3, 4),
  anchor = "dose",
  dose_optimization_method = map_adjust_dose
)
initial_method <- create_initial_regimen_design(
  method = weight_based_starting_dose,
  mgkg = 3.2,
  regimen = list(
    interval = 24,
    n = 4,
    t_inf = 3,
    type = "infusion"
  )
)
sim_mod_design <- create_model_design(lib = "pkbusulfanmccune")
est_mod_design <- create_model_design(lib = "pkbusulfanshukla")

design <- create_trial_design(
  sampling_design = tdm_design,
  target_design = target_design,
  regimen_update_design = update_design,
  initial_regimen_design = initial_method,
  sim_design = sim_mod_design, 
  est_design = est_mod_design
)
```

## 2. Create a set of digital patient covariates

We will use the NHANES data set for covariate distributions. This is a data set
of five thousand individuals collected by the US National Center for Health
Statistics. See package documentation for more details.

We will also use the `dplyr` package for some basic data processing. Since every
data set is un-tidy in its own way (“Tidy datasets are all alike, but every 
messy dataset is messy in its own way.” 
–– [Hadley Wickham](https://r4ds.had.co.nz/tidy-data.html)), data cleaning 
should be handled outside of the `mipdtrial` tools. The functions expect all
patient covariates to be numeric.

```{r}
dat <- NHANES::NHANES %>%
  # take just the first record per patient
  group_by(ID) %>%
  slice(1) %>%
  ungroup() %>%
  # convert columns to numeric
  mutate(Gender = ifelse(Gender == "male", 1, 0)) %>% # as defined in models
  select(ID, Gender, Age, Weight, Height) %>%
  # include only patients with all information available
  filter(!is.na(Gender), !is.na(Weight), !is.na(Height), !is.na(Age))
```

The next step is to make sure we have all the covariates we need. Busulfan has a
known time-dependent relationship with clearance, and so the McCune model takes
a covariate that indicates at what time this relationship should start
(T_CL_EFF; useful for test doses that should not impact busulfan clearance). We
will set this to zero for all patients. The Shukla model also takes conditioning
regimen as a covariate, since co-medication with Clofarabine and Fludarabine was
associated with a change in clearance (REGI). We will also set this to zero for
all patients.

Other models might require fat-free mass or other calculated covariates. This 
would be a good time to do that sort of processing on your data set!

You may also need to convert between units (height in m to cm, for example.)

```{r}
dat$T_CL_EFF <- 0
dat$REGI <- 0
```

Here are the first few rows of our data set:

```{r}
head(dat)
```

We will also only look at the first 10 patients in our data set, in the 
interest of speed.

```{r}
dat <- head(dat, 10)
```

We also need to link the covariates in our data set to the covariates expected
in the model.

```{r}
# check which covariates are required for your model using 
# `attr(sim_mod_design$model, "covariates")`

# check which covariates are in your data set using `colnames(dat)`
cov_map <- c(
  AGE = "Age", 
  WT = "Weight", 
  HT = "Height", 
  SEX = "Gender",
  T_CL_EFF = "T_CL_EFF",
  REGI = "REGI"
)

```


## 3. Simulate a trial!

Each patient will start at a dose of 3.2 mg/kg body weight. We will then use
`mipdtrial::dose_grid_search` to optimize doses to attain a target of 90 mg*h/L.

For inter-individual variability terms, we will use the inter-individual
variability described by our "true" model, since this should reflect the
"true" distribution in patient pharmacokinetics. Later, we will estimate
the individual PK parameters using our estimation model and the measured
TDMs we "collect".

For residual error terms, we will use the residual error of our "true" model,
since this is supposed to reflect the "true" unexplained error (assay error, 
etc.). This function creates a data frame of error to add to each "true"
simulated concentration to produced a measured therapeutic drug monitoring 
sample.

```{r}
res <- run_trial(data = dat, design = design, cov_mapping = cov_map, seed = 1)

```

Here are the first few rows of our simulation results:

```{r}
head(res$final_exposure)
```

## 6. Analyze results

How well did our patients get to target? It looks like we estimated that
target attainment would be very high (all patients have an AUC very close to 
90 mg*h/L), but "true" AUC was higher than that due to some model 
misspecification!

```{r fig.width = 2, fig.height = 3}
res$final_exposure %>%
  pivot_longer(
    c(auc_true, auc_est),
    names_to = "auc_type",
    values_to = "auc"
  ) %>%
  ggplot() +
    aes(x = auc_type, y = auc/1000) +
    geom_boxplot() +
    geom_hline(yintercept = 90, color = "navyblue", linetype = "dashed") +
    scale_y_continuous(limit = c(0, NA)) +
    theme_minimal() +
    labs(
      y = "Cumulative AUC (mg\u00B7h/L)",
      x = "AUC type"
    )
```

Even with some model misspecification, target attainment (within 20% of target 
AUC) was still high:

```{r}
target_attainment <- res$final_exposure %>%
  mutate(
    ontarget = is_on_target(auc_true, target_design)) %>%
  summarize(proportion_ontarget = mean(ontarget)) %>%
  pull(proportion_ontarget)
target_attainment <- paste0(round(100 * target_attainment), "%")
```

Overall, `r target_attainment` of patients had a "true" AUC of 90 mg*h/L.

For an in-depth analysis of how differences between these two models impact
target attainment, see [Hughes et al., *J PKPD* (2024)](https://link.springer.com/article/10.1007/s10928-024-09915-w)

