---
title: "Model misspecification, a busulfan case study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Busulfan MIPD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, 
  message = FALSE
)
```

In this vignette, we will demonstrate the following tools:

* using {mipdtrial} to simulate a trial, with the trial design specified using
  R code.
* simulating patient PK and estimating patient PK with different models, to
  assess the impact of model misspecification
* using a weight-based initial dose

To account for model misspecification, we will use one model to simulate
(McCune, 2014) and one model to estimate (Shukla, 2020). This is a simplified
version of a [published
study](https://link.springer.com/article/10.1007/s10928-024-09915-w) in which we
compared non-compartmental analysis (NCA) with MAP Bayesian estimation for
AUC-guided busulfan dosing.

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# for CI, most windows users should not need this since the .libPaths should
# include where the model was installed.
if (.Platform$OS.type == "windows") { 
    .libPaths(c("D:/a/mipdtrial/mipdtrial/check/mipdtrial.Rcheck", .libPaths()))
}
if(!requireNamespace("pkbusulfanmccune", quietly = TRUE)) {
  capture.output(PKPDsim::install_default_literature_model("pk_busulfan_mccune"))
}
if (!requireNamespace("pkbusulfanshukla", quietly = TRUE)) {
  capture.output(capture.output(PKPDsim::install_default_literature_model("pk_busulfan_shukla")))
}
```

```{r setup, message=FALSE, warning=FALSE}
library(mipdtrial)
if(!requireNamespace("pkbusulfanmccune", quietly = TRUE)) {
  PKPDsim::install_default_literature_model("pk_busulfan_mccune")
  loadNamespace("pkbusulfanmccune")
}
if (!requireNamespace("pkbusulfanshukla")) {
  PKPDsim::install_default_literature_model("pk_busulfan_shukla")
  loadNamespace("pkbusulfanshukla")
}

# For data handling/plotting
library(dplyr)
library(tidyr)
library(ggplot2)

```

## 1. Define trial design

In this simple trial, patients will be dosed once per day for four days, with 4
levels collected per day, at 3.5, 4, 6 and 8 hours post-dose. We will adjust
doses 2, 3, and 4 to achieve a target cumulative AUC of 90 by t = 192 hours.

```{r}
tdm_design <- create_sampling_design(
  offset = rep(c(3.5, 4, 6, 8), 4),
  at = rep(1:4, each = 4),
  anchor = "dose"
)
target_design <- create_target_design(
  targettype = "cum_auc", 
  targetvalue = 90 * 1000,       # unit conversion
  time = 192
)
update_design <- create_regimen_update_design(
  at = c(2, 3, 4),
  anchor = "dose",
  dose_optimization_method = map_adjust_dose
)
initial_method <- create_initial_regimen_design(
  method = weight_based_starting_dose,
  mgkg = 3.2,
  dosing_weight = "Weight", # matches column in data set
  regimen = list(
    interval = 24,
    n = 4,
    t_inf = 3,
    type = "infusion"
  )
)
sim_mod_design <- create_model_design(lib = "pkbusulfanshukla")
est_mod_design <- create_model_design(lib = "pkbusulfanmccune")

design <- create_trial_design(
  sampling_design = tdm_design,
  target_design = target_design,
  regimen_update_design = update_design,
  initial_regimen_design = initial_method,
  sim_design = sim_mod_design, 
  est_design = est_mod_design
)
```

## 2. Create a set of digital patient covariates

We will create an example data set of 10 individuals originally drawn from the
NHANES data set. This data set was created from survey data collected by the US
National Center for Health Statistics from a general sampling of the US
population, and is available in the NHANES R package.

```{r}
dat <- data.frame(
  ID = 1:10,
  Gender = c(1, 1, 0, 1, 1, 0, 1, 1, 1, 0),
  Age = c(34L, 4L, 49L, 9L, 8L, 45L, 66L, 58L, 54L, 10L),
  Weight = c(87.4, 17, 86.7, 29.8, 35.2, 75.7, 68, 78.4, 74.7, 38.6),
  Height = c(164.7, 105.4, 168.4, 133, 130.6, 166.7, 169.5, 182, 169.4, 141.8)
)
```

The next step is to make sure we have all the covariates we need. Busulfan has a
known time-dependent relationship with clearance, and so the McCune model takes
a covariate that indicates at what time this relationship should start
(T_CL_EFF; useful for test doses that should not impact busulfan clearance). We
will set this to zero for all patients. The Shukla model also takes conditioning
regimen as a covariate, since co-medication with Clofarabine and Fludarabine was
associated with a change in clearance (REGI). We will also set this to zero for
all patients.

Other models might require fat-free mass or other calculated covariates. This 
would be a good time to do that sort of processing on your data set!

You may also need to convert between units (height in m to cm, for example.)

```{r}
dat$T_CL_EFF <- 0
dat$REGI <- 0
```

Here are the first few rows of our data set:

```{r}
head(dat)
```

We also need to link the covariates in our data set to the covariates expected
in the model:

  - To check which covariates are required for your models use
    `PKPDsim::get_model_covariates()`:

    ```{r}
    unique(
      c(
        PKPDsim::get_model_covariates(sim_mod_design$model),
        PKPDsim::get_model_covariates(est_mod_design$model)
      )
    )
    ```
  
  - To check which covariates are in your data set use `colnames()`:

    ```{r}
    colnames(dat)
    ```


```{r}
cov_map <- c(
  AGE = "Age", 
  WT = "Weight", 
  HT = "Height", 
  SEX = "Gender",
  T_CL_EFF = "T_CL_EFF",
  REGI = "REGI"
)
```


## 3. Simulate a trial!

Each patient will start at a dose of 3.2 mg/kg body weight. We will then use
`mipdtrial::dose_grid_search` to optimize doses to attain a target of 90 mg*h/L.

For inter-individual variability terms, we will use the inter-individual
variability described by our "true" model, since this should reflect the
"true" distribution in patient pharmacokinetics. Later, we will estimate
the individual PK parameters using our estimation model and the measured
TDMs we "collect".

For residual error terms, we will use the residual error of our "true" model,
since this is supposed to reflect the "true" unexplained error (assay error, 
etc.). This function creates a data frame of error to add to each "true"
simulated concentration to produced a measured therapeutic drug monitoring 
sample.

```{r echo = FALSE}
res <- run_trial(data = dat, design = design, cov_mapping = cov_map, seed = 1, progress = FALSE)
```

Here are the first few rows of our simulation results:

```{r}
head(res$final_exposure)
```

## 6. Analyze results

How well did our patients get to target? It looks like we estimated that
target attainment would be very high (all patients have an AUC very close to 
90 mg*h/L), but "true" AUC was higher than that due to some model 
misspecification!

```{r fig.width = 2, fig.height = 3}
res$final_exposure %>%
  pivot_longer(
    c(auc_true, auc_est),
    names_to = "auc_type",
    values_to = "auc"
  ) %>%
  ggplot() +
    aes(x = auc_type, y = auc/1000) +
    geom_boxplot() +
    geom_hline(yintercept = 90, color = "navyblue", linetype = "dashed") +
    scale_y_continuous(limit = c(0, NA)) +
    theme_minimal() +
    labs(
      y = "Cumulative AUC (mg\u00B7h/L)",
      x = "AUC type"
    )
```

Even with some model misspecification, target attainment (within 20% of target 
AUC) was still high:

```{r}
target_attainment <- res$final_exposure %>%
  mutate(
    ontarget = is_on_target(auc_true, target_design)) %>%
  summarize(proportion_ontarget = mean(ontarget)) %>%
  pull(proportion_ontarget)
target_attainment <- paste0(round(100 * target_attainment), "%")
```

Overall, `r target_attainment` of patients had a "true" AUC of 90 mg*h/L.

For an in-depth analysis of how differences between these two models impact
target attainment, see [Hughes et al., *J PKPD* (2024)](https://link.springer.com/article/10.1007/s10928-024-09915-w)

