---
title: "Busulfan MIPD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Busulfan MIPD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this example script, we will estimate target attainment for busulfan.

To account for model misspecification, we will use one model to simulate 
(McCune, 2014) and one model to estimate (Shukla, 2020).

```{r setup}
library(mipdtrial)
require(pkbusulfanmccune)
require(pkbusulfanshukla)

# For example data
library(NHANES)

# For data handling/plotting
library(dplyr)
library(tidyr)
library(ggplot2)

```


## 1. Create a set of digital patient covariates

We will use the NHANES data set for covariate distributions. This is a data set
of five thousand individuals collected by the US National Center for Health
Statistics. See package documentation for more details.

We will also use the `dplyr` package for some basic data processing. Since every
data set is un-tidy in its own way (“Tidy datasets are all alike, but every 
messy dataset is messy in its own way.” 
–– [Hadley Wickham](https://r4ds.had.co.nz/tidy-data.html)), data cleaning 
should be handled outside of the `mipdtrial` tools. The functions expect all
patient covariates to be numeric.

```{r}
dat <- NHANES::NHANES %>%
  # take just the first record per patient
  group_by(ID) %>%
  slice(1) %>%
  ungroup() %>%
  # convert columns to numeric
  mutate(Gender = ifelse(Gender == "male", 1, 0)) %>% # as defined in models
  select(ID, Gender, Age, Weight, Height) %>%
  # include only patients with all information available
  filter(!is.na(Gender), !is.na(Weight), !is.na(Height), !is.na(Age))
```

The next step is to make sure we have all the covariates we need. Busulfan has a
known time-dependent relationship with clearance, and so the McCune model takes
a covariate that indicates at what time this relationship should start
(T_CL_EFF; useful for test doses that should not impact busulfan clearance). We
will set this to zero for all patients. The Shukla model also takes conditioning
regimen as a covariate, since co-medication with Clofarabine and Fludarabine was
associated with a change in clearance (REGI). We will also set this to zero for
all patients.

Other models might require fat-free mass or other calculated covariates. This 
would be a good time to do that sort of processing on your data set!

You may also need to convert between units (height in m to cm, for example.)

```{r}
dat$T_CL_EFF <- 0
dat$REGI <- 0
```

Here are the first few rows of our data set:

```{r}
head(dat)
```

## 2. Define simulation parameters

In this simple trial, patients will be dosed once per day, with 4 levels
collected per day, at 3.5, 4, 6 and 8 hours post-dose. We are aiming for a
target cumulative AUC of 90 over the course of t = 0 hours to t = 192 hours.

```{r}
tdm_times <- rep(c(3.5, 4, 6, 8), 4) + rep(c(0, 24, 48, 72), each = 4)

target <- list(
  type = "cum_auc",
  value = 90 * 1000 # unit conversion
)
auc_times <- c(0, 192)

```

We will also only look at the first 10 patients in our data set, in the 
interest of speed.

```{r}
ids <- dat$ID[1:10]

```

## 3. Load model definitions

We need to define the model parameters and variability terms for our simulation.
For this study, we will use the ones defined in the models.

```{r}
# Set up simulation model ("true")
model_sim <- pkbusulfanmccune::model()
parameters_sim <- pkbusulfanmccune::parameters()
omega_sim <- pkbusulfanmccune::omega_matrix()
ruv_sim <- pkbusulfanmccune::ruv()

# Set up estimation model (model for MIPD)
model_est <- pkbusulfanshukla::model()
parameters_est <- pkbusulfanshukla::parameters()
omega_est <- pkbusulfanshukla::omega_matrix()
ruv_est <- pkbusulfanshukla::ruv()
```

We also need to link the covariates in our data set to the covariates expected
in the model.

```{r}
# check which covariates are required for your model using 
# `attr(model_sim, "covariates")`

# check which covariates are in your data set using `colnames(dat)`
cov_mapping <- c(
  AGE = "Age", 
  WT = "Weight", 
  HT = "Height", 
  SEX = "Gender",
  T_CL_EFF = "T_CL_EFF",
  REGI = "REGI"
)

```

## 4. Pre-generate variability terms

Reproducible analyses are crucial. Pre-generating residual error terms and
inter-individual variability terms allow you to easily return to an analysis
without worrying about seed position.

For residual error terms, we will use the residual error of our "true" model,
since this is supposed to reflect the "true" unexplained error (assay error, 
etc.). This function creates a data frame of error to add to each "true"
simulated concentration to produced a measured therapeutic drug monitoring 
sample.

```{r}
res_var <- pregenerate_ruv(
  ids = ids,
  n_iter = 1, # just one copy of each patient for now
  tdm_sample_time = tdm_times,
  prop = ruv_sim$prop,
  add = ruv_sim$add,
  seed = 60
)

```

For inter-individual variability terms, we will use the inter-individual
variability described by our "true" model, since this should reflect the
"true" distribution in patient pharmacokinetics. Later, we will estimate
the individual PK parameters using our estimation model and the measured
TDMs we "collect".

```{r}
pkpar <- pregenerate_iiv(
  ids = ids,
  n_iter = 1, # just one copy of each patient for now
  sim_model = model_sim,
  omega = omega_sim,
  parameters = parameters_sim,
  seed = 60
)
```

For an extended analysis, it would be prudent to save these data frames as
CSV or RDS files for easier accessibility.

## 5. Initialize output metrics of interest

We want to record both "true" cumulative AUC and estimated cumulative AUC, 
along with patient identifiers (i.e., id). We will pre-populate a data frame
rather than iteratively binding rows, since that is more performent.

```{r}
results <- data.frame(
  id = ids,
  iter = 1,
  true_auc = NA_real_,
  est_auc = NA_real_
)
```

## 6. Simulate a trial!

Each patient will start at a dose of 3.2 mg/kg body weight. We will then use
`mipdtrial::dose_grid_search` to optimize doses to attain a target of 90 mg*h/L.

```{r}
for (i in ids) {
  # get patient covariates
  covs <- create_cov_object(
    dat[dat$ID == i,],
    mapping = cov_mapping
  )
  # create initial dosing regimen (see PKPDsim::new_regimen for more info)
  reg <- PKPDsim::new_regimen(
    amt = 3.2 * covs$WT$value, # we will update this each "day" of the trial.
    interval = 24,
    n = 4,
    t_inf = 3,
    type = "infusion"
  )
  ruv_i <- res_var[res_var$id == i,]
  par_true_i <- as.list(pkpar[pkpar$id == i,])
  # initialize objects for loop
  tdms_i <- data.frame()
  
  # MIPD trial
  for (day in 1:4) {
    # collect TDMs from today (use model for simulation!)
    today <- (tdm_times >= (day - 1) * 24 & tdm_times < day * 24)
    new_tdms <- collect_tdms(
      sim_model = model_sim,
      t_obs = tdm_times[today],
      res_var = ruv_i[today,],
      pars_i = par_true_i,
      regimen = reg,
      covariates = covs
    )
    tdms_i <- bind_rows(tdms_i, new_tdms)
    
    # get MAP fit, using model for estimation
    est_par <- simulate_fit(
      est_model = model_est,
      parameters = parameters_est,
      omega = omega_est,
      ruv = ruv_est,
      tdms = tdms_i,
      covariates = covs,
      regimen = reg
    )
    
    # calculate new dose, using the estimation model, if we aren't on last day
    if (day < 4) {
      new_dose <- dose_grid_search(
        est_model = model_est,
        regimen = reg,
        parameters = est_par, # we want to use our "best guess" to get the dose
        t_obs = auc_times[2],
        target = target,
        obs_comp = attr(model_est, "size"), # last compartment is AUC
        dose_update = day + 1, # update dose starting from next day
        dose_grid = seq(from = 1, to = 400, by = (400 - 1) / 10),
        covariates = covs,
        iov_bins = attr(model_est, "iov")$bins
      )
      reg <- update_regimen(reg, new_dose, dose_update_number = day + 1)
    }
  }
    
  # get final "true" and "estimated" cumulative AUC at end
  sim_true <- PKPDsim::sim(
    model_sim,
    parameters = par_true_i, # true patient parameters
    regimen = reg,
    covariates = covs,
    t_obs = auc_times,
    iov_bins = attr(model_sim, "iov")$bins
  )
  auc_true <- calc_auc_from_sim(sim_true, attr(model_sim, "size"))
  sim_est <- PKPDsim::sim(
    model_est,
    parameters = est_par, # best guess at patient parameter
    regimen = reg,
    covariates = covs,
    t_obs = auc_times,
    iov_bins = attr(model_est, "iov")$bins
  )
  auc_est <- calc_auc_from_sim(sim_est, attr(model_est, "size"))
  
  results$true_auc[results$id == i] <- auc_true
  results$est_auc[results$id == i] <- auc_est
}

```

Here are the first few rows of our simulation results:

```{r}
head(results)
```

## 7. Analyze results

How well did our patients get to target? It looks like we estimated that
target attainment would be very high (all patients have an AUC very close to 
90 mg*h/L), but "true" AUC was higher than that due to some model 
misspecification!

```{r fig.width = 2, fig.height = 3}
results %>%
  pivot_longer(
    c(true_auc, est_auc),
    names_to = "auc_type",
    values_to = "auc"
  ) %>%
  ggplot() +
  aes(x = auc_type, y = auc/1000) +
  geom_boxplot() +
  geom_hline(yintercept = 90, color = "navyblue", linetype = "dashed") +
  scale_y_continuous(limit = c(0, NA)) +
  theme_minimal() +
  labs(
    y = "Cumulative AUC (mg\u00B7h/L)",
    x = "AUC type"
  )

```

