---
title: "PEG Asparaginase"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PEG Asparaginase}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, 
  message = FALSE
)
```

```{r echo=FALSE}
# for CI, most windows users should not need this since the .libPaths should
# include where the model was installed.
if (.Platform$OS.type == "windows") { 
    .libPaths(c("D:/a/mipdtrial/mipdtrial/check/mipdtrial.Rcheck", .libPaths()))
}
if(!require(pkpegasparaginasemodifiedwurthwein, warn.conflicts = FALSE)) {
  PKPDsim::install_default_literature_model("pk_pegasparaginase_modified_wurthwein")
}
```


```{r setup}
library(mipdtrial)
library(PKPDsim) # for working with models
if(!require(pkpegasparaginasemodifiedwurthwein, warn.conflicts = FALSE)) {
  install_default_literature_model("pk_pegasparaginase_modified_wurthwein")
}
library(NHANES) # for example data

# For data handling/plotting
library(dplyr, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)

```

## 1. Define simulation parameters

PEG-asparaginase has a long half-life. In this simulation trial patients 
will receive 4 doses separated by 14 days levels collected every 14 days just 
prior to the next dose. We will adjust doses 2 and 3 to achieve a steady state 
asparaginase activity level of of 300 IU/L (0.3 IU/mL) (goal: 0.1 - 0.5 IU/mL)

```{r}
tdm_times <- c( 13.9*24,  27.9*24,  41.9*24) 
adjust_at_dose <- c(2, 3) 

target <- create_target_object(
  targettype = "conc", 
  targetvalue = 300
)

conc_times <- 1344
```

## 2. Create a set of digital patient covariates

Synthetic dataset based on CDC growth curves of children.

```{r}
dat <- NHANES::NHANES %>%
  # take just the first record per patient
  group_by(ID) %>%
  slice(1) %>%
  ungroup() %>%
  # convert columns to numeric
  mutate(Sex = ifelse(Gender == "male", 1, 0)) %>% # as defined in models
  filter(Age<18) %>%
  select(ID, Sex, Age, Weight, Height) %>%
  # include only patients with all information available
  filter(!is.na(Sex), !is.na(Weight), !is.na(Height), !is.na(Age)) %>%
  mutate(BSA = sqrt(Weight * Height / 3600))
dat <- na.omit(dat)
head(dat)
```

Here are the first few rows of our data set:

```{r}
head(dat)
```

While our data set is much larger, for this example let's take just the first
20 patients. 

```{r}
ids <- sample(dat$ID,20)
```

## 3. Load model definitions

We need to define the model parameters and variability terms for our simulation.
For this study, we will use the ones defined in the models.

```{r}
# Set up simulation model 
model <- pkpegasparaginasemodifiedwurthwein::model()
parameters <- pkpegasparaginasemodifiedwurthwein::parameters()
omega <- pkpegasparaginasemodifiedwurthwein::omega_matrix()
ruv <- pkpegasparaginasemodifiedwurthwein::ruv()
```

We also need to link the covariates in our data set to the covariates expected
in the model.

```{r}
# check which covariates are required for your model using 
# `attr(model, "covariates")`

# check which covariates are in your data set using `colnames(dat)`
cov_mapping <- c(
  AGE = "Age", 
  WT = "Weight", 
  HT = "Height", 
  BSA = "BSA",
  SEX = "Sex"
)
```

## 4. Initialize output metrics of interest

We will pre-populate a data frame rather than iteratively binding rows, since
that is more performent.

```{r}
results <- data.frame(
  id = rep(ids,each=3),
  iter = c(1,2,3),
  day = c(14,28,42),
  amt = rep(NA_real_,each=3),
  conc = rep(NA_real_,each=3)
)
```


## 5. Simulate a trial!

Patients will start at a dose of 2000 IU/m2 if <22 yo and 2500 IU/m2 if >22 yo, 
with doses capped at a maximum total dose of 3750 IU. We will then use
`mipdtrial::dose_grid_search` to optimize doses to attain a target concentration 
of 0.3 IU/mL.

For inter-individual variability terms, we will use the inter-individual
variability described by our model, since this should reflect the distribution 
in patient pharmacokinetics. Later, we will estimate the individual PK 
parameters using our estimation model and the measured TDMs we "collect".

For residual error terms, we will use the residual error of our  model,
since this is supposed to reflect the unexplained error (assay error, 
etc.). This function creates a data frame of error to add to each simulated 
concentrations to produced a measured therapeutic drug monitoring sample.

```{r}
set.seed(1) # important for reproducibility

for (i in ids) {
  # get patient covariates
  covs <- create_cov_object(
    dat[dat$ID == i,],
    mapping = cov_mapping
  )
  # use age-based nomogram
  dose <- ifelse(covs$AGE$value >= 22, 2500*covs$BSA$value, 2000*covs$BSA$value)
  # capt dose to a maximum of 3750
  dose <- pmin(dose, 3750)
  
  # create initial dosing regimen (see PKPDsim::new_regimen for more info)
  reg <- new_regimen(
    amt = dose, 
    interval = 14*24,
    n = 4,
    t_inf = 1,
    type = "infusion"
  )
  
  # randomly draw individual PK parameters
  pars_true_i <- generate_iiv(
    sim_model = model,
    omega = omega,
    parameters = parameters
  )
  
  treatment_summary <- sample_and_adjust_by_dose(
    adjust_at_dose = adjust_at_dose,
    tdm_times = tdm_times,
    regimen = reg,
    covariates = covs,
    pars_true_i = pars_true_i,
    sim_model = model,
    sim_ruv = ruv,
    est_model = model,
    parameters = parameters,
    omega = omega,
    ruv = ruv,
    target_time = conc_times,
    target = target,
    dose_optimization_method = dose_adjust_map
  )
    
  # get final asparaginase activity level
  sim <- PKPDsim::sim(
    model,
    parameters = as.list(pars_true_i), # true patient parameters
    regimen = treatment_summary$final_regimen,
    covariates = covs,
    ruv = ruv,
    t_obs = c(14, 28, 42) * 24,
    only_obs = TRUE
  )
  conc  <- sim$y
  results$conc[results$id == i] <- conc
  
  #Add amounts
  amt <- treatment_summary$final_regimen$dose_amts
  results$amt[results$id == i] <- amt[-1]
}
```

Here are the first few rows of our simulation results:

```{r}
head(results)
```

## 6. Analyze results

How well did our patients get to target?

```{r}
results %>%
  filter(day %in% c(14,28,42)) %>%
  filter(!is.na(conc)) %>%
  mutate(
    below_flag = conc < 100,
    within_flag = conc >= 100 & conc <= 500,
    above_flag = conc > 500
  ) %>%
  group_by(day) %>%
  summarize(
    below_per = mean(below_flag),
    within_per = mean(within_flag),
    above_per = mean(above_flag)
  ) %>%
  ungroup() %>%
  pivot_longer(
    c(below_per,within_per,above_per),
    names_to = "per_type",
    values_to = "per"
  ) %>%
  mutate(
    per_type = case_when(
      per_type == "below_per" ~ "C. < 0.1 IU/mL",
      per_type == "within_per" ~ "B. 0.1 - 0.5 IU/mL",
      per_type == "above_per" ~ "A. > 0.5 IU/mL"
    )
  ) %>%
  select(day, per, per_type) %>%
  ggplot(
    aes(
      x = day,
      y = per*100,
      group = per_type,
      fill = per_type,
      label = ifelse(round(per * 100,1) == 0, "" , round(per*100, 1)))
  ) +
  geom_bar(stat="identity") +
  geom_text(size = 3, position = position_stack(vjust = 0.6), color="black") +
  theme_minimal() +
  ylab("Percent (%)") +
  xlab("Time (d)") +
  theme(
    legend.title = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  scale_x_continuous(breaks = c(14,28,42), labels = c(14,28,42)) +
  scale_fill_manual(values = c("#3870FA", "#35BCB1", "#FF8C00"))
```

While further work is needed to establish a realistic upper toxicity threshold
to inform the therapeutic window of PEG-asparaginase, an MIPD approach informed
by AAL levels which are already germane to therapeutic care with
PEG-asparaginase may translate to better target attainment and cost savings. For
further details, see [Brooks et al., PAGE Meeting; June 26-28, 2024, Rome,
Italy. (https://www.page-meeting.org/default.asp?abstract=10796)].
