---
title: "Specifying a trial using yaml, a sample timing case study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Effect of sample timing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r echo=FALSE, results='hide'}
# for CI, most windows users should not need this since the .libPaths should
# include where the model was installed.
if (.Platform$OS.type == "windows") { 
    .libPaths(c("D:/a/mipdtrial/mipdtrial/check/mipdtrial.Rcheck", .libPaths()))
}

if (!require(pkvancothomson)) {
  PKPDsim::install_default_literature_model("pk_vanco_thomson")
}
```

In this vignette, we will demonstrate the following tools:

- Using {mipdtrial} to simulate a trial, with the trial design specified using
  yaml files.
- We will have two trial "arms", varying only by the therapeutic drug monitoring
  strategy used. The two arms will be defined by two yaml files.

## Motivation

A hospital currently collects two samples for adjusting vancomycin doses: one
at 1-hr post-dose and another at 9 hours post-dose. They want to compare the 
impact on target attainment (an area under the curve (AUC) of 400-600 mg*h/L) if
they were to switch to collecting a single sample at 5 hours post-dose.

Here is how we could answer that problem using simulation!

```{r setup,message=FALSE}
library(mipdtrial)
library(dplyr)   # for easier data manipulation 
library(tidyr)
library(ggplot2) # for plotting our results
```

## 1. Define trial design

This simulated trial will have two arms: 

1. Two samples, collected 1 and 9 hours after a dose
2. One sample, collected 5 hours after a dose

In each case, we will collect the samples in the fourth dosing interval, and for simplicity, we will assume all patients are receiving vancomycin twice daily, infused over 2 hours.

We will adjust the fifth dose based on these levels, aiming for a daily AUC of  400-600 mg*h/L by day 6.

These designs are implemented in YAML files supplied with the package as follows:

```{r, echo=TRUE, eval=FALSE}
name: Test different sampling designs.
description: Sampling design 1 - peak / trough
designs:
  initial_regimen:
    method: model_based_starting_dose
    regimen:
      interval: 12
      type: infusion
      t_inf: 1
    settings:
      auc_comp: 3
      dose_resolution: 250
      dose_grid: [250, 5000, 250]
  target:
    targettype: auc24
    targetmin: 400
    targetmax: 600
    at: 6
    anchor: day
  sampling:
    offset: [1, 9]
    at: [4, 4]
    anchor: dose
  regimen_update:
    at: [5]
    anchor: dose
    update_type: dose
    dose_optimization_method: map_adjust_dose
  sim:
    lib: pkvancothomson
  est:
    lib: pkvancothomson
```

We can read this spec into R using `create_trial_design()`:

```{r}
arm1 <- create_trial_design(
  file = system.file(package = "mipdtrial", "md/sample_timing1.yaml")
)
arm2 <- create_trial_design(
  file = system.file(package = "mipdtrial", "md/sample_timing2.yaml")
)
```

## 2. Create a set of digital patient covariates

For this example, we will randomly generate a set of weights and creatinine
clearances (CRCLs) for our synthetic data set.

```{r}
set.seed(15)
dat <- data.frame(
  ID = 1:30,
  weight = rnorm(30, 90, 25),               # kg, normally distributed
  crcl = exp(rnorm(30, log(6), log(1.6)))   # L/hr, log-normally distributed 
)
```

We will use the Thomson (2009) model, which accepts additional clearance from
hemodialysis as a covariate. Let's set that to zero in our data set. 

Other models might require fat-free mass or other calculated covariates. This 
would be a good time to do that sort of processing on your data set!

```{r}
dat$CL_HEMO <- 0
```

Here are the first few rows of our data set:

```{r}
head(dat)
```

We need to link the covariates in our data set to the covariates expected
in the model.
```{r}
cov_mapping <- c(
  WT = "weight", 
  CRCL = "crcl",
  CL_HEMO = "CL_HEMO"
)
```

## 3. Simulate a trial!

Patients will get a model-based dose (using population PK parameters), and then
this dose will be adjusted based on the MAP Bayesian fit made using the 
collected samples.

Individual PK parameters will be randomly generated based on the interindividual
variability described in the model, and residual variability will be added to
each sample collected using the error model described in the model.

```{r}
res1 <- run_trial(
  design = arm1,
  data = dat,
  cov_mapping = cov_mapping,
  seed = 15,
  progress = FALSE
)
res2 <- run_trial(
  design = arm2,
  data = dat,
  cov_mapping = cov_mapping,
  seed = 15,
  progress = FALSE
)
```

## 4. Analyze results

We are interested in AUC target attainment. How did target attainment compare
between the two arms of the trial?

```{r}
results <- bind_rows(
  res1$final_exposure %>% mutate(sampling_design = "peak-trough"),
  res2$final_exposure %>% mutate(sampling_design = "mid-interval"),
)
target_attainment <- results %>%
  mutate(ontarget = ifelse(auc_true >= 400 & auc_true <= 600, 1, 0)) %>%
  group_by(sampling_design) %>%
  summarize(prop_on_target = 100 * mean(ontarget))
target_attainment %>%
  ggplot() +
    aes(x = sampling_design, y = prop_on_target) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    theme(
      panel.grid.major.x = element_blank()
    ) +
    labs(
      x = "Sampling strategy",
      y = "Percent on target (%)"
    )
```

Target attainment was high and varied little between the two arms, providing
evidence to support the move from collecting two samples to collecting a single
sample per dosing interval.

Because we are simulating each sampling strategy in each patient, we can also
look at how each patient responded to each sampling strategy.

```{r}
data.frame(
  peak_trough = res1$final_exposure$auc_true,
  random = res2$final_exposure$auc_true
) %>%
  ggplot() +
    aes(x = peak_trough, y = random) +
    geom_rect(
      aes(xmin = 400, xmax = 600, ymin = -Inf, ymax = Inf), 
      fill = "grey70",
      alpha = 0.05
    ) +
    geom_rect(
      aes(ymin = 400, ymax = 600, xmin = -Inf, xmax = Inf),
      fill = "grey70",
      alpha = 0.05
    ) +
    geom_point() +
    theme_minimal()
```

There is a strong correlation in final AUC between the two sampling strategies.
Some patients were under-exposed or over-exposed in both strategies, while 
others were on-target in one strategy but not in the other.
