---
title: "Effect of sample timing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Effect of sample timing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A hospital currently collects two samples for adjusting vancomycin doses: one
at 1-hr post-dose and another at 9 hours post-dose. They want to compare the 
impact on target attainment (an area under the curve (AUC) of 400-600 mg*h/L) if
they were to switch to collecting a single sample at 5 hours post-dose.

Here is how we could answer that problem using simulation!

```{r setup,message=FALSE}
library(mipdtrial)
library(dplyr)   # for easier data manipulation 
library(tidyr)
library(ggplot2) # for plotting our results
```

## 1. Define trial design

This simulated trial will have two arms: 

1. Two samples, collected 1 and 9 hours after a dose
2. One sample, collected 5 hours after a dose

In each case, we will collect the samples in the fourth dosing interval, and for simplicity, we will assume all patients are receiving vancomycin twice daily,
infused over 2 hours.

```{r}
# end_of_fourth_infusion <- (4-1) * 12 + 2 # first dose at t=0, infusion is 2 hrs
# tdm_times1 <- c(1, 9) + end_of_fourth_infusion
# tdm_times2 <- 5 + end_of_fourth_infusion
tdm_design1 <- create_sampling_design(
  time = c(1, 9),
  offset_from = c("dose", "dose"), 
  anchor = c(4, 4), 
  anchor_by = "dose"
)
tdm_design2 <- create_sampling_design(
  time = 5, 
  offset_from = "dose", 
  anchor = 4,
  anchor_by = "dose"
)
```

We will adjust the fifth dose based on these levels, aiming for a daily AUC of 
400-600 mg*h/L by day 6.

```{r}
update_design <- create_regimen_update_design(
  anchor = 5,
  anchor_by = "dose"
)
target_design <- create_target_design(
  targettype = "auc24", 
  targetmin = 400,
  targetmax = 600,
  anchor = 6,
  anchor_by = "day"
)
```

## 2. Create a set of digital patient covariates

For this example, we will randomly generate a set of weights and creatinine
clearances (CRCLs) for our synthetic data set.

```{r}
set.seed(15)
dat <- data.frame(
  ID = 1:30,
  weight = rnorm(30, 90, 25),               # kg, normally distributed
  crcl = exp(rnorm(30, log(6), log(1.6)))   # L/hr, log-normally distributed 
)
```

We will use the Thomson (2009) model, which accepts additional clearance from
hemodialysis as a covariate. Let's set that to zero in our data set. 

Other models might require fat-free mass or other calculated covariates. This 
would be a good time to do that sort of processing on your data set!

```{r}
dat$CL_HEMO <- 0
```

Here are the first few rows of our data set:

```{r}
head(dat)
```

## 3. Load model definitions

We need to define the model parameters and variability terms for our simulation.
For this study, we will use the ones defined in the models.


```{r echo=FALSE, results='hide'}
# for CI, most windows users should not need this since the .libPaths should
# include where the model was installed.
if (.Platform$OS.type == "windows") { 
    .libPaths(c("D:/a/mipdtrial/mipdtrial/check/mipdtrial.Rcheck", .libPaths()))
}

if (!require(pkvancothomson)) {
  PKPDsim::install_default_literature_model("pk_vanco_thomson")
}
```

```{r message=FALSE, warning=FALSE}
# Set up simulation model
if (!require(pkvancothomson)) {
  PKPDsim::install_default_literature_model("pk_vanco_thomson")
  library(pkvancothomson)
}
model <- pkvancothomson::model()
pop_parameters <- pkvancothomson::parameters()
omega <- pkvancothomson::omega_matrix()
ruv <- pkvancothomson::ruv()
```

We also need to link the covariates in our data set to the covariates expected
in the model.

```{r}
# check which covariates are required for your model using 
# `PKPDsim::get_model_covariates(model)`

# check which covariates are in your data set using `colnames(dat)`
cov_mapping <- c(
  WT = "weight", 
  CRCL = "crcl",
  CL_HEMO = "CL_HEMO"
)

```

## 4. Initialize output metrics of interest

We want to compare AUC at the end of the treatment course according to each 
sample collection strategy. We will pre-populate a data frame rather than 
iteratively binding rows, since that is more performant.

```{r}
results <- data.frame(
  id = dat$ID,
  peak_trough = NA_real_,
  random = NA_real_
)
```

## 5. Simulate a trial!

Patients will get a model-based dose (using population PK parameters), and then
this dose will be adjusted based on the MAP Bayesian fit made using the 
collected samples.

Individual PK parameters will be randomly generated based on the interindividual
variability described in the model, and residual variability will be added to
each sample collected using the error model described in the model.

```{r}
set.seed(15) # important for reproducibility

for (i in dat$ID) {
  # get patient covariates
  covs <- create_cov_object(
    dat[dat$ID == i,],
    mapping = cov_mapping
  )
  # find initial starting dose: define basic regimen, then upda
  initial_reg <- model_based_starting_dose(
    target_design = target_design,
    sampling_design = tdm_design1,
    n = 12,
    interval = 12,
    t_inf = 2,
    dose_resolution = 250, # round to nearest 250 mg
    grid = seq(250, 6000, by = 250),
    grid_type = "dose",
    # define model components
    est_model = model,
    parameters = pop_parameters,
    covariates = covs,
    auc_comp = attr(model, "size")
  )

  # randomly draw individual PK parameters
  pars_true_i <- generate_iiv(
    sim_model = model,
    omega = omega,
    parameters = pop_parameters
  )
  
  # Summary of both MIPD approaches
  treatment_summary1 <- sample_and_adjust_by_dose(
    regimen_update_design = update_design, 
    sampling_design = tdm_design1, # arm 1
    target_design = target_design,
    regimen = initial_reg,
    covariates = covs,
    pars_true_i = pars_true_i,
    sim_model = model,
    sim_ruv = ruv,
    est_model = model,
    parameters = pop_parameters,
    omega = omega,
    ruv = ruv,
    dose_optimization_method = map_adjust_dose
  )
  treatment_summary2 <- sample_and_adjust_by_dose(
    regimen_update_design = update_design, 
    sampling_design = tdm_design2, # arm 1
    target_design = target_design,
    regimen = initial_reg,
    covariates = covs,
    pars_true_i = pars_true_i,
    sim_model = model,
    sim_ruv = ruv,
    est_model = model,
    parameters = pop_parameters,
    omega = omega,
    ruv = ruv,
    dose_optimization_method = map_adjust_dose
  )
  
  # get steady-state AUC at end of treatment course
  results$peak_trough[results$id == i] <- calc_auc_from_regimen(
    regimen = treatment_summary1$final_regimen,
    parameters = pars_true_i, # true patient parameters
    model = model,
    target_design = target_design,
    covariates = covs
  )
  results$random[results$id == i] <- calc_auc_from_regimen(
    regimen = treatment_summary2$final_regimen,
    parameters = pars_true_i, # true patient parameters
    model = model,
    target_design = target_design,
    covariates = covs
  )
}
```

Here are the first few rows of our simulation results:

```{r}
head(results)
```

## 6. Analyze results

We are interested in AUC target attainment. How did target attainment compare
between the two arms of the trial?

```{r}
target_attainment <- results %>%
  pivot_longer(
    c(peak_trough, random),
    names_to = "sampling_design",
    values_to = "steady_state_auc"
  ) %>%
  mutate(ontarget = is_on_target(steady_state_auc, target_design)) %>%
  group_by(sampling_design) %>%
  summarize(prop_on_target = 100 * mean(ontarget))

target_attainment %>%
  ggplot() +
    aes(x = sampling_design, y = prop_on_target) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    theme(
      panel.grid.major.x = element_blank()
    ) +
    labs(
      x = "Sampling strategy",
      y = "Percent on target (%)"
    )
```

Target attainment was high and varied little between the two arms, providing
evidence to support the move from collecting two samples to collecting a single
sample per dosing interval.

Because we are simulating each sampling strategy in each patient, we can also
look at how each patient responded to each sampling strategy.

```{r}
results %>%
  ggplot() +
    aes(x = peak_trough, y = random) +
    geom_rect(
      aes(xmin = 400, xmax = 600, ymin = -Inf, ymax = Inf), 
      fill = "grey70",
      alpha = 0.05
    ) +
    geom_rect(
      aes(ymin = 400, ymax = 600, xmin = -Inf, xmax = Inf),
      fill = "grey70",
      alpha = 0.05
    ) +
    geom_point() +
    theme_minimal()
```

There is a strong correlation in final AUC between the two sampling strategies.
Some patients were under-exposed or over-exposed in both strategies, while 
others were on-target in one strategy but not in the other.
