---
title: "Dose and interval adaptation (vancomycin)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dose and interval adaptation (vancomycin)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In model-informed precision dosing we can update doses, but of course we can also reduce or increase the dosing interval length. This can also be studied using 
the `mipdtrial` package.

```{r setup}
library(mipdtrial)
library(dplyr)   # for easier data manipulation 
library(tidyr)
library(ggplot2) # for plotting our results
```

## 1. Define the trial design

This simulated trial will have two samples, with a peak and trough collected
in dose 1 and 3.

```{r}
# end_of_fourth_infusion <- (4-1) * 12 + 2 # first dose at t=0, infusion is 2 hrs
# tdm_times1 <- c(1, 9) + end_of_fourth_infusion
# tdm_times2 <- 5 + end_of_fourth_infusion
tdm_design1 <- create_sampling_design(
  time = c(1, -1, 1, -1), # sample 1-hour before next dose, and at peak
  offset_from = c("peak", "trough", "peak", "trough"),
  anchor = c(1, 1, 3, 3),
  anchor_by = "dose"
)
```

We assume that we can update the dose amount or dosing interval
at dose 3 and 5 and aim for an AUC24 of 400-600 mg*h/L by day 6.

```{r}
target_design <- create_target_design(
  targettype = "auc24", 
  targetmin = 400,
  targetmax = 600,
  time = 0,
  anchor = 6,
  anchor_by = "day"
)
dose_update_design <- create_regimen_update_design(
  anchor = c(3, 5),
  anchor_by = "dose",
  update_type = "dose"
)
```

## 2. Create a set of digital patient covariates

For this example, we will randomly generate a set of weights and creatinine
clearances (CRCLs) for our synthetic data set. See `sampling_timing()` vignette
for a longer description.

```{r}
set.seed(15)
dat <- data.frame(
  ID = 1:30,
  weight = rnorm(30, 90, 25),               # kg, normally distributed
  crcl = exp(rnorm(30, log(6), log(1.6)))   # L/hr, log-normally distributed 
)
dat$CL_HEMO <- 0
```

## 3. Load model definitions

We need to define the model parameters and variability terms for our simulation.
For this study, we will use the ones defined in the models.

```{r echo=FALSE}
# Set up simulation model
if (!require(pkvancothomson)) {
  PKPDsim::install_default_literature_model("pk_vanco_thomson")
  library(pkvancothomson)
}
model <- pkvancothomson::model()
pop_parameters <- pkvancothomson::parameters()
omega <- pkvancothomson::omega_matrix()
ruv <- pkvancothomson::ruv()
```

We also need to link the covariates in our data set to the covariates expected
in the model.

```{r}
# check which covariates are required for your model using 
# `PKPDsim::get_model_covariates(model)`

# check which covariates are in your data set using `colnames(dat)`
cov_mapping <- c(
  WT = "weight", 
  CRCL = "crcl",
  CL_HEMO = "CL_HEMO"
)
```

## 4. Simulate a trial!

Patients will get a model-based dose (using population PK parameters), and then
we will adjust either the dose or the interval based on the MAP Bayesian fit made
using the collected samples.

Individual PK parameters will be randomly generated based on the interindividual
variability described in the model, and residual variability will be added to
each sample collected using the error model described in the model.

First let's simulate the "dose optimization" scenario, leaving the interval
fixed.

```{r}
set.seed(15) # important for reproducibility
tdms <- c()
dose_updates <- c()
for (i in dat$ID) {
  
  #################################################################################
  ## Create individual
  #################################################################################
  # get patient covariates
  covs <- create_cov_object(
    dat[dat$ID == i,],
    mapping = cov_mapping
  )
  # randomly draw individual PK parameters
  pars_true_i <- generate_iiv(
    sim_model = model,
    omega = omega,
    parameters = pop_parameters
  )

  #################################################################################
  ## Find initial starting dose
  #################################################################################
  # find initial starting dose: define basic regimen, then upda
  initial_reg <- model_based_starting_dose(
    sampling_design = tdm_design1,
    target_design = target_design,
    n = 12,
    interval = 12,
    t_inf = 2,
    dose_resolution = 250, # round to nearest 250 mg
    grid = seq(250, 6000, by = 250),
    grid_type = "dose",
    est_model = model,
    parameters = pop_parameters,
    covariates = covs,
    auc_comp = attr(model, "size")
  )

  #################################################################################
  ## Main patient-level loop: run through regimen optimization
  #################################################################################
  treatment_summary <- sample_and_adjust_by_dose(
    regimen_update_design = dose_update_design, 
    sampling_design = tdm_design1, # arm 1
    target_design = target_design,
    regimen = initial_reg,
    covariates = covs,
    pars_true_i = pars_true_i,
    sim_model = model,
    sim_ruv = ruv,
    est_model = model,
    parameters = pop_parameters,
    omega = omega,
    ruv = ruv,
    target_time = auc_time,
    dose_optimization_method = map_adjust_dose
  )
  
  #################################################################################
  ## Collect data into object
  #################################################################################
  tdms <- bind_rows(
    tdms,
    treatment_summary$tdms %>% mutate(id = i)
  )
  dose_updates <- bind_rows(
    dose_updates,
    treatment_summary$dose_updates %>% mutate(id = i)
  )
  
}
```

## 5. Analyze results for dose-adaptation

Let's see how well that worked. In the figure below we plot the steady state
AUC that is attained using the regimen updated at the time shown on the x-axis.
You would expect over time the steady state AUC to approach the middle of the
range more closely, as more TDM is available to optimize the dose. Since
we're keeping the interval fixed, each patient will have the same dose update
times on the x-axis.

```{r}
ggplot(dose_updates) +
  geom_line(aes(x = t, y = auc_before_update, group = id)) +
  geom_point(aes(x = t, y = auc_before_update, group = id)) +
  geom_ribbon(data = data.frame(t = c(0, max(dose_updates$t)), ymin = c(400, 400), ymax = c(600, 600)), aes(x = t, ymin = ymin, ymax = ymax), fill = "blue", alpha = 0.2) +
  xlab("Weeks") + ylab("Trough concentration (mg/L)")
```

## 6. Interval adaptation

Next, we'll simulate a trial in which the *interval* is optimized, and the
dose is fixed (after an initial model-based dose is obtained). Again, we'll
allow the dose to be changed after doses 3 and 5 (irrespective of interval).

```{r}
interval_update_design <- create_regimen_update_design(
  anchor = c(3, 5),
  anchor_by = "dose",
  update_type = "interval"
)
```

```{r}
set.seed(15) # important for reproducibility
tdms <- c()
dose_updates <- c()
for (i in dat$ID) {
  
  #################################################################################
  ## Create individual
  #################################################################################
  # get patient covariates
  covs <- create_cov_object(
    dat[dat$ID == i,],
    mapping = cov_mapping
  )
  # randomly draw individual PK parameters
  pars_true_i <- generate_iiv(
    sim_model = model,
    omega = omega,
    parameters = pop_parameters
  )

  #################################################################################
  ## Find initial starting dose
  #################################################################################
  # find initial starting dose: define basic regimen, then upda
  initial_reg <- model_based_starting_dose(
    sampling_design = tdm_design1,
    target_design = target_design,
    n = 30,
    interval = 12,
    t_inf = 2,
    dose_resolution = 250, # round to nearest 250 mg
    grid = seq(250, 6000, by = 250),
    grid_type = "dose",
    # define model components
    est_model = model,
    parameters = pop_parameters,
    covariates = covs,
    auc_comp = attr(model, "size")
  )

  #################################################################################
  ## Main patient-level loop: run through regimen optimization
  #################################################################################
  treatment_summary <- sample_and_adjust_by_dose(
    regimen_update_design = interval_update_design,
    sampling_design = tdm_design1, # arm 1
    target_design = target_design,
    regimen = initial_reg,
    covariates = covs,
    pars_true_i = pars_true_i,
    sim_model = model,
    sim_ruv = ruv,
    est_model = model,
    parameters = pop_parameters,
    omega = omega,
    ruv = ruv,
    target_time = auc_time,
    dose_optimization_method = map_adjust_interval,
    grid = c(6, 8, 12, 16, 24, 36),
    verbose = F
  )
  
  #################################################################################
  ## Collect data into object
  #################################################################################
  # extract metrics of interest
  tdms <- bind_rows(
    tdms,
    treatment_summary$tdms %>% mutate(id = i)
  )
  dose_updates <- bind_rows(
    dose_updates,
    treatment_summary$dose_updates %>% mutate(id = i)
  )
}
```

## 7. Analyze results for interval-adaptation

When we plot the results of this simulation, we now see that dose update times
are flexible, and not the same for each patient. We again see that the resulting
steady state AUC24 also approaches the middle of the targeted AUC range. So this
shows that the dosing interval can also be used very easily to obtain optimal 
drug exposure.

```{r}
ggplot(dose_updates) +
  geom_line(aes(x = t, y = auc_before_update, group = id)) +
  geom_point(aes(x = t, y = auc_before_update, group = id)) +
  geom_ribbon(data = data.frame(t = c(0, max(dose_updates$t)), ymin = c(400, 400), ymax = c(600, 600)), aes(x = t, ymin = ymin, ymax = ymax), fill = "blue", alpha = 0.2) +
  xlab("Time of regimen update") + ylab("Resulting AUC at t=6 days")
```

```{r}
dose_updates %>%
  mutate(idx = rep(c(1:3), 30)) %>%
  ggplot() +
    geom_bar(aes(x = as.factor(interval_before_update))) +
    facet_wrap(~idx) +
    xlab("Interval")
```
