---
title: "Time-varying exposure targets (vancomycin)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time-varying exposure targets (vancomycin)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r echo=FALSE, results='hide'}
# for CI, most windows users should not need this since the .libPaths should
# include where the model was installed.
if (.Platform$OS.type == "windows") { 
    .libPaths(c("D:/a/mipdtrial/mipdtrial/check/mipdtrial.Rcheck", .libPaths()))
}

if(!requireNamespace("pkvancothomson", quietly = TRUE)) {
  PKPDsim::install_default_literature_model("pk_vanco_thomson")
  loadNamespace("pkvancothomson")
}
```

In this vignette, we will demonstrate the following tools:

- using {mipdtrial} to simulate a trial, with the trial design specified using
  R code.
- using timevarying exposure targets in the trial simulations. This is not
  so relevant for vancomycin, but we'll just use that here as an example drug.

```{r setup}
library(mipdtrial)
library(dplyr)   # for easier data manipulation 
library(tidyr)
library(ggplot2) # for plotting our results
```

## 1. Define the trial design

This simulated trial will have two samples, with a peak and trough collected
in dose 1 and 3.

```{r}
tdm_design <- create_sampling_design(
  offset = c(1, -1, 1, -1), # sample 1-hour before true trough, and at peak+1hr
  when = c("peak", "trough", "peak", "trough"),
  at = c(1, 1, 3, 3),
  anchor = "dose"
)
```

We assume that we can update the dose amount at dose 3 and 7, and aim for an 
AUC24 of 500 mg*h/L and 300 by dose 5 and 10, respectively. \

```{r}
target_design <- create_target_design(
  targettype = "auc24", 
  targetvalue = c(500, 300),
  at = c(5, 10),
  anchor = "dose"
)
dose_update_design <- create_regimen_update_design(
  at = c(3, 7),
  anchor = "dose",
  update_type = "dose",
  dose_optimization_method = map_adjust_dose # fixed interval, optimize dose
)
```

We will be using the Thomson (2009) model for simulation and estimation:

```{r}
model_design <- create_model_design(lib = "pkvancothomson")
```

We will start with a dose estimated to attain the target exposure metrics based
on population PK parameters, assuming a 12-hour interval.

```{r}
initial_method <- create_initial_regimen_design(
  method = model_based_starting_dose,
  regimen = list(
    interval = 12,
    type = "infusion",
    t_inf = 1,
    n = 24
  ),
  settings = list(
    auc_comp = 3,
    dose_resolution = 250,
    dose_grid = c(250, 5000, 250)
  )
)
```

Now we can combine these design choices into a single trial design:

```{r}
design <- create_trial_design(
  sampling_design = tdm_design, 
  target_design = target_design,
  regimen_update_design = dose_update_design, # arm 1
  initial_regimen_design = initial_method,
  sim_design = model_design, est_design = model_design
)
```

## 2. Create a set of digital patient covariates

For this example, we will randomly generate a set of weights and creatinine
clearances (CRCLs) for our synthetic data set. See `sampling_timing()` vignette
for a longer description.

```{r}
set.seed(15)
dat <- data.frame(
  ID = 1:30,
  weight = rnorm(30, 90, 25),               # kg, normally distributed
  crcl = exp(rnorm(30, log(6), log(1.6)))   # L/hr, log-normally distributed 
)
dat$CL_HEMO <- 0 # required covariate for our model
```

We also need to link the covariates in our data set to the covariates expected
in the model:

  - To check which covariates are required for your model use
    `PKPDsim::get_model_covariates()`:

    ```{r}
    PKPDsim::get_model_covariates(model_design$model)
    ```
  
  - To check which covariates are in your data set use `colnames()`:

    ```{r}
    colnames(dat)
    ```

```{r}
cov_map <- c(
  WT = "weight", 
  CRCL = "crcl",
  CL_HEMO = "CL_HEMO"
)
```

## 4. Simulate a trial!

Patients will get a model-based dose (using population PK parameters), and then
we will adjust the dose based on the MAP Bayesian fit made
using the collected samples, first to target an AUC of 500, and then later 
targeting an AUC of 300.

Individual PK parameters will be randomly generated based on the interindividual
variability described in the model, and residual variability will be added to
each sample collected using the error model described in the model.

```{r}
covs <- create_cov_object(
  dat,
  mapping = cov_map
)
model_based_starting_dose(
  design = design,
  covariates = covs,
  cov_mapping = cov_mapping
)
res <- run_trial(
  data = dat,
  design = design,
  cov_mapping = cov_map,
  progress = FALSE,
  seed = 15
)
```
## 5. Analyze results 

Let's see how well that worked. In the figure below we plot the steady state
AUC that is attained using the regimen updated at the time shown on the x-axis.
You would expect over time the steady state AUC to approach the middle of the
range more closely, as more TDM is available to optimize the dose. Since
we're keeping the interval fixed, each patient will have the same dose update
times on the x-axis.

```{r fig.width = 5}
ggplot(res$dose_updates) +
  aes(x = t, y = auc_before_update, group = id) +
  geom_hline(yintercept = c(300, 500), linetype = "dotted") +
  geom_line() +
  geom_point() +
  xlab("Time of regimen update (Dose update algorithm)") + 
  ylab("Resulting AUC at time points initial dose, and dose 3, 5 (mg*h/L)")
```
